---
title: "Elastic Net"
output: html_document
---

Our professor introduced a new method or should I say it more formally, a new regularization and variable selection method know as Elastic Net. For those people wondering what regularization is, it was a bit of a head scratcher for me at the beginning too. Let me give you guys a brief idea.

**Regularization** : Regularization is the process of introducing additional information to solve problems presented by overfitting.

This method seems to outperfrom lasso while it still enjoys the similar form of sparsity we see in lasso.

##Introduction

To evaluate the qualiy of any model, two aspects are pretty important.

1) Accuracy of predicting new data - I must say as Statisticians, this is one area that is immensely powerful to us and doing it well is really important

2) Interpretation of the model - This is yet another important area because a simpler model is definitely easier to interpret.

##Penalization techniques

Penalization techniques have been known to improve methods related to OLS. Some examples of penalization techniques are lasso and ridge regression. Looking at the advantages and the disadvantages is a good way to revise the topic.

**Lasso**

Advantage : Due to the nature of the $L_1$ penalty, the lasso does continuous shrinkage and automatic variable selection. As mentioned earlier, in terms of interpretabililty, using lasso will be beneficial.

Disadvantages : If we have a group of variables among which we have high correlations, lasso tends to select one variable from that group and it does not bother which one is chosen.

**Ridge**

Advantages: Ridge regression achieves better prediction performance through a bias-variance trade off. To those people wondering what in the world bias-variance trade off is, I will give a small explanation at the end of this page.

Disadvantages: Ridge regression cannot produce a parsimonious model because it keeps all the predictors in the model.

##Elastic Net

Based on the explanation given above, we can see that both lasso and ridge regression have their advantages and disadvantages. However, it was a thought by Hastie and Zou that it would be interesting to find a method that works as well as lasso and fix other issues seen by lasso. And out of this thought came about Elastic Net.

###Properties of Elastic Net

1) The elastic net simultaneously does variable selection and continuous shrinkage
2) It is able to select a group of variables

###Properties of Naive Elastic Net

Even above starting to fully understand about elastic net, I did not know there was a naive elastic net. Let me give a brief explanation.

Let **y**=$(y_1,y_2,....y_n)^t$ be the response and **x**=$[x_1,x_2,...x_p]$ be the explanatory variables. Here we note that $x_1 = (x_{1j},x_{2j}.....,x_{3j})^t$, j=1,2....p are predictors.

Using the above, we can define the naive elastic net as a combination of the L1 and L2 penalty. 

$\hat\beta(Naive Enet) = arg min_\beta ||y-x\beta||^2 + \lambda_1|\beta_1| + \lambda_2||\beta||^2$

Limitations : The two-stage procedure for naive elastic net incurs double amount of shrinkage and introduces extra bias without reducing variance

###Properties of Elastic Net

$\hat\beta$(elastic net) = $(1+\lambda_2)\hat\beta$(naive elastic net)

##Comparison of the Elastic Net and Naive Elastic Net

I tried different $\lambda_2$ values to do a comparison between the elastic net and naive elastic net. My results indicate that for small values of $\lambda_2$, the prediction error for corrected elastic net was smaller than the prediction error for naïve elastic net.

```{r echo = FALSE}
naive = c(16.95229, 16.94592, 16.94389, 16.94324)
elastic = c(14.39191, 16.49244, 16.67565, 16.8162)
lambda2 <- c(0.0001, 0.00001, 0.000001, 0.0000001)
a <- rbind(lambda2,naive,elastic)
```

```{r echo = FALSE}
library(knitr)
kable(round(a))
```

